---
title: "hdarticle17range: Distribution and range calculation tool for Habitats Directive Article 17 reporting"
bibliography: references.bib
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Package 'hdarticle17range' implements several functions that simplify distribution and range calculation of species and habitat types in various member states and biogeographical regions.

Range of a species or habitat type is a crucial parameter in Article 17 reporing under Habitats Directive. While distribution is a representation of the occurrences in the 10x10 km grid, range calculation is more sophisticated. Range is obtained by generalizing the distribution, by creating envelope(s) around the distribution grids using a parameter named 'gap distance' [@guidelines].

#### Datasets

| Dataset           | Description                                                                       |
| :---------------- | :-------------------------------------------------------------------------------- |
| eu10km            | EEA reference grid (10 km) geometries [@eearefgrid].                              |
| biogeoregions10km | EEA refecence grid (10 km) of biogeographic regions [@biogeoregions,@msfdregions].|
| ms10kmcells       | EEA reference grids (10 km) of member states [@eearefgrid].                       |
| gapdistance       | Default gap distance for species and habitats.                                    |
| eesti             | Simplified country borders for Estonia.                                           |

### Functions

| Function                     | Description                                                                                 |
| :--------------------------- | :------------------------------------------------------------------------------------------ |
| ms\_region\_reference_grid() | Prepare a reference related to specific member state and biogeographic region.              |
| features\_to\_range()        | Calculate distribution and range from input occurence data, reference grid and gap distance.|

## Installation

You can install the development version from [GitHub](https://github.com/) with:

```{r installpackage, eval = FALSE}
# install.packages("devtools")
devtools::install_github("leivimee/hdarticle17range")
```

## Usage

```{r loadpackages, eval=TRUE, results='hide'}
# load packages
library(hdarticle17range)
```

Grid cells of a biogeographic regions in specific member state (MS) can be obtained with a function called 'ms_region_reference_grid'. The function takes two important arguments -- 'memberstate' and 'biogeoregion'. For example, to retrieve the reference grid for Estonia in Boreal biogeographic region, one can call:

```{r example1, eval = TRUE, include = TRUE, fig.width=6, fig.height=4}
# prepare cells
eebor10km <- ms_region_reference_grid("EE", "BOR")
# load ggplot2 for graphic output
library(ggplot2)
# plot
ggplot(data = eebor10km)+
  geom_sf()+
  theme_void()
```

The function makes a query from grid cells that have intersection with both -- with the EEA given MS grid cells and EEA given BG regions. Sometimes it is needed to exclude some cells that are common with some neighbouring MS. It is possible to fine tune the reference grid by providing country borders.

```{r example2, eval = TRUE, include = TRUE, fig.width=6, fig.height=4}
# load example dataset
data(eesti)
# prepare cells
eebor10kmee <- ms_region_reference_grid("EE", "BOR", eesti)
# plot
ggplot(data = eebor10kmee)+
  geom_sf()+
  theme_void()
```

Lets calculate distribution and range maps from dummy data.

```{r example3, eval = TRUE, include = TRUE}
# set seed for reproducible example
set.seed(69)
# generate 20 random points
exa_spe_occ <- sf::st_as_sf(sf::st_sample(eesti, 20))
exa_spe_range <- features_to_range(exa_spe_occ, eebor10kmee)
```

Examine the structure of output.

```{r example4, eval = TRUE, include = TRUE}
str(exa_spe_range)
```

The output consists of a list with elements **dist** and **range**. Element **dist** contains a list with three elements: 

**dist**
: geometry (sfc POLYGON) of grid cells

**area** 
: area (units) of distribution in units [km^2]

**cells**
: a vector (character) containing cell codes


Element **range** contains a list with three elements: 

**range**
: geometry (sfc POLYGON) of grid cells

**area**
: area (units) of range in units [km^2]

**cells**
: a vector (character) containing cell codes

Plot the range object.

```{r example5, eval = TRUE, include = TRUE, fig.width=6, fig.height=4}
ggplot(data = exa_spe_range$range$range)+
  geom_sf()+
  geom_sf(data=eesti, fill=NA, color="#cccccc")+
  geom_sf(data=exa_spe_occ)+
  theme_void()
```



## References

